### 2024-10-03 texts/paper_closing.tex
more re-writing/re-ordering of hypotheses

% Created 2024-10-02 wo 15:08
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

@@

In recent years, private collector-run museums have seen an unprecedented proliferation, with hundreds of organizations founded since the beginning of the millenium \parencite{Velthuis_etal_2023_boom,LarrysList_2015_report}.
These museums, run by individual collectors, constitute an addition to established organizational forms of art provision such as museums operated and financed by the state, foundations or associations.
While many aspects of this new organizational form has attracted attention, such as their emergence, relations to other institutions or impact on artist consecration (cf. \cite{Kolbe_etal_2022_privatemuseum} for a literature review), the process of closure has received considerably less attention. 
Furthermore, the focus that has been given to questions of sustainability has been primarily conducted through case studies; quantitative studies have been mostly absent.
In this paper, I address this gap by investigating the closure event of members of this new organizational form.
Analysis which factors are associated with museum closures provides insight not only into the event of closing, but also into the ongoing processes of a private museums, such as dependency relationships.
Furthermore, analyzing which museums survive allows to make informed speculations on the future of this recently emerged organizational form.

@@

Due to the founders central role, private museums have been argued to be highly dependent on their founder for their continuing existence, as museums generally cannot cover expenses via revenues from ticket prices, museum shops and cafes:
The \textcite{IMLS_2008_funding} found that American non-profit art museums receive less than half (48.1\%) of their revenues from earned income, leaving substantial gaps in the budgets that are filled by a mix of private donations (23.0\%), investment (18.5\%) and government support (10.4\%) (p.27).
Outside of the United States, the reliance on revenue streams beyond operations is even larger:
German museums in the 1990s covered on average only around 30\% of their budget by their income \parencite{Martin_1993_museen}\footnote{Own calculations based on table p.233}, with private museums however requiring less support as operations provide around 47\% of their budget (ibid.).
In the German cultural sector, non-profit organizations more widely  draw around 50\% of their funding from public support \parencite[p.82]{Zimmer_Priller_2007_gemeinnuetzig}.
On a European level, only 12\% of museum incomes are generated by entry fees, whereas 69\% originate from subsidies \parencite{EGMUS_2024_complete}.\footnote{Own calculations from EGMUS data, based on all countries with income data from 2015 onwards.}

Given that operating at "the margin of financial sustainability" is the "ordinary condition" for cultural NPOs, \parencite[p.2]{Licci_BaraldiBonini_2024_sustainability}, it is not surprising that costs have been listed as a primary reason for private museum closure:
\textcite{Adam_2021_rise} argues that "most of the cases of defunct private museums come back to the issue of cost" (p.79); a sentiment shared by \textcite{Walker_2019_collector} who notes that "private institutions struggle to contain escalating costs that are associated with running a museum" (p.234).
In other studies, "insufficient funding, high maintenance costs and lack of strong government support" \parencite[p.7]{Zennaro_2017_shanghai}, a "shortage of fundings and interruptions in cash" \parencite[p.45]{Song_2008_private}, bankruptcies \parencite{Velthuis_Gera_2024_fragility,Liu_2019_identities,DeNigris_2018_museums}, costly legal battles \parencite{Velthuis_Gera_2024_fragility}, stock devaluations \parencite{Walker_2019_collector} or failing auction sales \parencite{Bechtler_Imhof_2018_future} were central reasons that forced private museums to close.
In a systematic review of closing reasons given in newspapers, magazines, social media accounts and personal exchanges, \textcite{Velthuis_Gera_2024_fragility} find financial issues as the most commonly mentioned reasons for private museum closing.
\subsubsection*{diversification}

@@

Existing research on private museum is conducted primarily by scholars of the art field or market as well as art historians.
Nethertheless, the identification of the centrality of the founder as a critical source of vulnerability corresponds strongly to the finding in non-profit management of a relation between diversification and mortality \parencite{Tuckman_Chang_1991_vulnerability}.
For example, \cite{Bielefeld_1994_survival} finds that nonprofits with a larger number of funding sources (e.g. from foundations, individuals, federated funders, corporations, trusts, government at federal, state, county and municipal level as well as self-generated sources such as dues, interests, and service revenue, p.31) are more likely to survive.
Other studies have measured diversification as the Herfindahl Index (a measure of concentration) of income sources such as public support, program service revenues, dues and assessments, net fundraising income and profits from sale of inventory \parencite[p.381]{Hager_2001_vulnerability} or contributions, program service revenue, investment income, other revenues \parencite[p.37]{Lu_Shon_Zhang_2019_dissolution}, with both studies finding a positive association between diversification and survival; in the case of \textcite{Hager_2001_vulnerability} also specifically for art museums.
Next to diversification, volatility, i.e. fluctuation in revenue streams, has been associated with mortality as well \parencite{Mayer_2022_simmer}.
Despite limited reception of this research stream in the private museum literature, identifying the high reliance on the founder as a source of risk thus corresponds to management insights. 
Resulting from the dependence of the museum on the founder, it is plausible that interruptions to this key revenue source would endanger the viability of the museum, which can be formulated as: 
\textbf{Hypothesis 1}: Events which disrupt the founder's ability to provide resources to their private museums increase the risk of closing.

The most direct interruption a founder and their museum can experience is the death of the founder; this event has in fact already been extensively argued to result in museum closure:
\textcite{Walker_2019_collector} argues that "the death of the original founder and creator can [\ldots{}] place the future of private museums and collections in jeopardy" as "seldom do heirs share a similar passion or wish to take on the financial burden of maintaining private museums indefinitely" (p.234). 
\textcite{Bechtler_Imhof_2018_future} argue that in order to ensure private museum longevity, "flexible, easily adapted designs for private museums" are necessary "to reduce the financial burden on their heirs" (p.188).
\textcite{Adam_2021_rise} argues that succession introduces financial strains as "collectors’ successors may not look kindly on what may well be a money pit when [\ldots{}] they will have to pick up the bills" (p.77), and also highlights the aesthetic dimension as "the nature of contemporary art is to be fresh and new, but what is ‘fresh and new’ for the founder might seem ‘old hat’ to their heirs". (p.76).
\textcite{Velthuis_Gera_2024_fragility} similarly argue that "the founder’s death will almost inevitably have a direct impact on the financial health of the museum" (p.10).
\textcite{StylianouLambert_etal_2014_museums}, after discussing a case of a private museum whose owner "does not believe that anyone else could love the museum as much as she does", argue that this "generates questions regarding the cultural sustainability of small private museums in cases where their initiators are not able to care for them any longer" (p.580).
The death of the founder thus constitutes a clear financial and organizational disruption for the museum, and as previous research has seen it as a precursor of museum closure, a hypothesis can be formulated as:
\textbf{Hypothesis 1a}: Private museums are more likely to close after the death of the founder.

A less obvious yet possibly impactful event that can jeopardize a founder's ability to sustain their museum and potentially shorten its lifespan is a decline of status of the collector within the art world.
Recognition bestowed by prominent art world authorities confers the acknowledgment of "cultural expertise" \parencite[p.1486]{Braden_2016_recognition}, thus validating the collector-founder's refined taste, which constitutes a form symbolic capital \parencite{Bourdieu_1993_production}.
This symbolic capital might provide opportunities to acquire material advantages such as collaborations with other institutions and access to broader audiences, ultimately bolstering the museum's longevity.

While limited research exists that investigates reputation effects for (art) museums, its effects have been investigated for non-profits more widely.
Here reputation is often referred to as an organization's resource or asset \parencite{Kong_Farrell_2010_image,Goldfarb_Zavyalova_Pillai_2018_Rao,Walker_McCarthy_2010_survival,Schloderer_Sarstedt_Ringle_2014_reputation}. 
Studies primarily focus on the effect of reputation on some tangible benefits such as donations and contributions \parencite{Sarstedt_2010_reputation,Mews_Boenigk_2012_blood,Heller_2008_alliances,Meijer_2009_reputation}, or its direct effect on survial \parencite{Lu_Shon_Zhang_2019_dissolution,Singh_1991_change,Hager_2001_vulnerability} \footnote{Furthermore, the term "survival" is sometimes used more loosely to refer to a non-profit's effective operation and financial condition rather than the more narrowly defined dissolution (e.g. \cite[p.58]{Santos_Laureano_Moro_2019_reputation}, \cite[p.15]{Heller_2008_alliances}, \cite[p.111]{Schloderer_Sarstedt_Ringle_2014_reputation})}, with most studies finding support for a positive effect of reputation.
Next to the reputation of the general public, studies have also found reputation effects among more narrowly defined groups, in particular peers \parencite{Padanyi_Gainer_2003_peerreputation} and networks of similar organizations on regional and national scale \parencite{Walker_McCarthy_2010_survival}.
In particular this reputation among more narrowly defined groups makes it likely that the reputation of the  \emph{founder} matters (beyond the reputation of the \emph{organization}, the focus on most prior research) given the highly personal character of the private museum initiative and the central role of the founder in negotiations with decision makers from other organizations or the public sector.

In the case of the founder losing his reputation, the consequences then might extend beyond personal demotivation.
Such a decline could also curtail potential opportunities for cooperation with other institutions and limit the museum's reach to larger audiences, thus posing a risk to the museum's continued operation and survival in the long term.
\textbf{Hypothesis 1b}: Private museums are more likely to close after a decline of their founder's status.

While the low diversification, i.e. the centrality of the founder for museum finances, has been argued to constitute the key feature of private museum vulnerability, it is also important to consider the other ways in which private museums, as (non-profit) organizations, are entangled with their environments.
Its ability to attract audiences, which may involve competition with other museums, secure cooperations with other museums, acquire sponsorships from cooperations and fundings from governments and other foundations possibly constitute factors that contribute to its viability and hence longevity.
While some of these aspects can be influenced by the founder (in particular the museum's location and strategies), other are less so.
In particular when they rely on the decision-making of other art-field actors, such as competition with other museums and sponsorship/funding choices, the influence of the individual founder is severely limited.
\paragraph*{Competition}

Private museums have been characterized as highly competitive.
While some self-descriptions focus on collaborative aspect, previous studies have highlighted how private museums compete with public museums over art works, audiences, sponsoring and donors \parencite[p.4]{Kolbe_etal_2022_privatemuseum}. 
So far the primary interest of investigating competition has been to evaluate whether private museums pose a threat to art provision by public institutions.
However, less focus has been given to the fact that operating in a competitive environment may also have consequences for the private museums themselves, in particular their survival aspects.
While private museums presumably depend on their founder to a large extent, income from tickets and facilities, as well sponsorships and public support possibly also constitute other non-negligible revenue streams.
While competition directly has not yet been investigated as a cause for private museum closing, a possible mechanism of competition, "insufficient interest from the public", has been identified as the second most mentioned closure reason by \textcite[p.6]{Velthuis_Gera_2024_fragility}. 
Lack of interest by the public can be a consequence of multiple processes.
While competition is one such processes, and applies if nearby competing institutions capture large portions of the audience, lack of interest can also result from a lack of audience (in the absence of competing institutions), which may be the case for private museums located in sparsely populated areas.
In both these cases, the resources that a private museum is able to receive from actors in its environment are limited.

\textbf{Hypothesis 3}: A private museum is more likely to close in a more competitive environment.
\paragraph*{Identity}

The mere availability of audiences and other potential donors might however not be sufficient if a museum is not perceived as deserving support; i.e. if its identity is perceived as inappropriate. 
An identity that is legitimate (understandable) and favorable (positive) has been argued to have a vast number of positive outcomes for organizations such as allowing higher profits, satisfaction, as well as weaker impact of shocks, and providing the opportunity to deviate from established patterns  \parencite{Lange_Lee_Dai_2010_reputation}. 
As gaining the approval of different stakeholders is necessary for being understood and/or supported by other organizations in the environment, having a clearly defined and favorable identity influences resource acquisition and thereby survival prospects \parencite{Rao_1994_reputation}.
Even for non-profit organizations such as museums, which do not produce tangible goods (a domain where identity-related effects are especially prevalent, e.g. \cite{Hsu_2015_granted,Bogaert_etal_2014_ecological}), a clear and positive identity might be helpful in a number of ways:
In the context of private museums, a negative identity could lead to lower visitor numbers, less discounts from art dealers, less government grants or lower chances on collaborations with other museums or corporations, resulting in lower revenues and higher acquisition costs.

The empirical record of the link between identity and survival for non-profit organizations however is mixed: 
For example \textcite{Bielefeld_1994_survival} finds that NPOs who pursue less legitimation strategies (obtaining endorsements, lobbying or contributing to local causes) are more likely to close, and \textcite{HernandezOrtiz_2022_discontinuity} observes a declining image or reputation as a cause of closure in a number of non-profits (p.115).
Nonetheless, other studies find little to no effect of legitimacy on survival, both when measured via density \parencite{Bogaert_etal_2014_ecological} or from archival sources and interviews \parencite{Fernandez_2007_dissolution}.
\textcite{Walker_McCarthy_2010_survival} find support for some, but not all legitimacy indicators.

@@

It is however necessary to consider that in particular the name of a museum might be associated with a museum's survival via identity, but also reflect founder intentions regarding the institutions longevity; this possibility is discussed in more detail in the results section.
\subsection*{Methods and Data}

A newly constructed database, assembled from a wide range of sources, is used to measure the survival of private museums (cf. \textcite{Velthuis_etal_2023_boom} for details).
Currently, the database includes 550 museums, of which 453 are currently open, 72 have closed and 25 have transformed into other organizational forms.\footnote{The 25 institutions which were private museums at some point and have since been transformed into other organizational forms are not included in this analysis and discussed in the limitations.}
Using this database, in particular year of opening and, if available, closing, it is possible to reconstruct the life course of the each institution with museum-year as the unit of analysis.
While the database is not limited to private museums of a particular time period, due to the recency of the private museum boom in the last two decades 80\% of the museums included in it have opened after 2000.
As the databse was constructed retrospectively from 2020 to 2022, I focus here on closing events from 2000 to 2021 to limit the possibility that some institutions were missed, as such risks are likely higher for earlier time periods.
Furthermore, a small number of museums are excluded, in particular 2 museums for which the closing date could not be found, 17 museums which opened in 2021 and later (as survival analyses requires positive survival times), 1 museum the opening year of which could not be determined, as well as 2 museums which closed before 2000 and thereby fall outside the period under investigation.
This leaves 503 private museums at risk which have accumulated 6096 museum-years and 68 closures.
This data allows to conduct survival analysis (also known as event history analysis), the most common statistical technique to investigate hazard rates and identify mortality-assocated covariates \parencite{Moore_2015_survival,Allison_2014_event}.

>>>>

% Created 2024-10-03 do 13:26
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

@@

In recent years, private collector-run museums have seen an unprecedented proliferation, with hundreds of organizations founded since the beginning of the millenium \parencite{Velthuis_etal_2023_boom,LarrysList_2015_report}.
These museums, run by individual collectors, constitute an addition to established organizational forms of art provision such as museums operated and financed by the state, foundations or associations.
While many aspects of this new organizational form has attracted attention, such as their emergence, relations to other institutions or impact on artist consecration (cf. \cite{Kolbe_etal_2022_privatemuseum} for a literature review), the process of closure has received considerably less attention. 
Furthermore, the focus that has been given to questions of sustainability has been primarily conducted through case studies; quantitative studies have been mostly absent.
In this paper, I address this gap by investigating the closure event of members of this new organizational form.
Analyzing which factors are associated with museum closures provides insight not only into the event of closing, but also into the ongoing processes of a private museums, such as dependency relationships.
Furthermore, analyzing which museums survive allows to make informed speculations on the future of this recently emerged organizational form.

@@

Due to the founders central role, private museums have been argued to be highly dependent on their founder for their continuing existence, as museums generally cannot cover expenses via revenues from ticket prices, museum shops and cafes:
The \textcite{IMLS_2008_funding} found that American non-profit art museums receive less than half (48.1\%) of their revenues from earned income, leaving substantial gaps in the budgets that are filled by a mix of private donations (23.0\%), investment (18.5\%) and government support (10.4\%) (p.27).
Outside of the United States, the reliance on revenue streams beyond operations appears even larger:
German museums in the 1990s covered on average only around 30\% of their budget by their income \parencite{Martin_1993_museen}\footnote{Own calculations based on table b), p.233}, with private museums however requiring less support as operations provide around 47\% of their budget (ibid.).
In the German cultural sector, non-profit organizations more widely draw around 50\% of their funding from public support \parencite[p.82]{Zimmer_Priller_2007_gemeinnuetzig}.
On a European level, only 12\% of museum incomes are generated by entry fees, whereas 69\% originate from subsidies \parencite{EGMUS_2024_complete}.\footnote{Own calculations from EGMUS data, based on all countries with income data from 2015 onwards.}

Given that operating at "the margin of financial sustainability" is the "ordinary condition" for cultural NPOs, \parencite[p.2]{Licci_BaraldiBonini_2024_sustainability}, it is not surprising that costs have been listed as a primary reason for private museum closure:
\textcite{Adam_2021_rise} argues that "most of the cases of defunct private museums come back to the issue of cost" (p.79); a sentiment shared by \textcite{Walker_2019_collector} who notes that "private institutions struggle to contain escalating costs that are associated with running a museum" (p.234).
In other studies, "insufficient funding, high maintenance costs and lack of strong government support" \parencite[p.7]{Zennaro_2017_shanghai}, a "shortage of fundings and interruptions in cash" \parencite[p.45]{Song_2008_private}, bankruptcies \parencite{Velthuis_Gera_2024_fragility,Liu_2019_identities,DeNigris_2018_museums}, costly legal battles \parencite{Velthuis_Gera_2024_fragility}, stock devaluations \parencite{Walker_2019_collector} or failing auction sales \parencite{Bechtler_Imhof_2018_future} have been found as central reasons that forced private museums to close.
In a systematic review of closing reasons given in newspapers, magazines, social media accounts and personal exchanges, \textcite{Velthuis_Gera_2024_fragility} find financial issues as the most commonly mentioned reasons for private museum closing.
\subsubsection*{diversification}

@@

Existing research on private museum is conducted primarily by scholars of the art field or market as well as art historians.
Nethertheless, the identification of the centrality of the founder as a critical source of vulnerability corresponds strongly to the finding in non-profit management of a relation between diversification and mortality \parencite{Tuckman_Chang_1991_vulnerability}.
For example, \textcite{Bielefeld_1994_survival} finds that nonprofits with a larger number of funding sources (e.g. from foundations, individuals, federated funders, corporations, trusts, government at federal, state, county and municipal level as well as self-generated sources such as dues, interests, and service revenue, p.31) are more likely to survive.
Other studies have measured diversification as the Herfindahl Index (a measure of concentration) of income sources such as public support, program service revenues, dues and assessments, net fundraising income and profits from sale of inventory \parencite[p.381]{Hager_2001_vulnerability} or contributions, program service revenue, investment income, other revenues \parencite[p.37]{Lu_Shon_Zhang_2019_dissolution}, with both studies finding a positive association between diversification and survival; in the case of \textcite{Hager_2001_vulnerability} also specifically for art museums.
Next to diversification, volatility, i.e. fluctuation in revenue streams, has been associated with mortality as well \parencite{Mayer_2022_simmer}.
Despite limited reception of this research stream in the private museum literature, identifying the high reliance on the founder as a source of risk thus corresponds to management insights. 
Resulting from the dependence of the museum on the founder, it is plausible that interruptions to this key revenue source would endanger the viability of the museum, which can be formulated as: 
\bigbreak
\noindent
\textbf{Hypothesis 1}: Events which disrupt the founder's ability to provide resources to their private museums increase the risk of closing.

The most direct interruption a founder and their museum can experience is the death of the founder; this event has in fact already been extensively argued to result in museum closure:
\textcite{Walker_2019_collector} argues that "the death of the original founder and creator can [\ldots{}] place the future of private museums and collections in jeopardy" as "seldom do heirs share a similar passion or wish to take on the financial burden of maintaining private museums indefinitely" (p.234). 
\textcite{Bechtler_Imhof_2018_future} argue that in order to ensure private museum longevity, "flexible, easily adapted designs for private museums" are necessary "to reduce the financial burden on their heirs" (p.188).
\textcite{Adam_2021_rise} argues that succession introduces financial strains as "collectors’ successors may not look kindly on what may well be a money pit when [\ldots{}] they will have to pick up the bills" (p.77), and also highlights the aesthetic dimension as "the nature of contemporary art is to be fresh and new, but what is ‘fresh and new’ for the founder might seem ‘old hat’ to their heirs". (p.76).
\textcite{Velthuis_Gera_2024_fragility} similarly argue that "the founder’s death will almost inevitably have a direct impact on the financial health of the museum" (p.10).
\textcite{StylianouLambert_etal_2014_museums}, after discussing a case of a private museum whose owner "does not believe that anyone else could love the museum as much as she does", argue that this "generates questions regarding the cultural sustainability of small private museums in cases where their initiators are not able to care for them any longer" (p.580).
The death of the founder thus constitutes a clear financial and organizational disruption for the museum, and as previous research has seen it as a precursor of museum closure, a hypothesis can be formulated as:
\bigbreak
\noindent
\textbf{Hypothesis 1a}: Private museums are more likely to close after the death of the founder.

A less obvious yet possibly impactful event that might jeopardize a founder's ability to sustain their museum and shorten its lifespan is a decline of status of the collector within the art world.
Recognition bestowed by prominent art world authorities confers the acknowledgment of "cultural expertise" \parencite[p.1486]{Braden_2016_recognition}, thus validating the collector-founder's refined taste, which constitutes a form symbolic capital \parencite{Bourdieu_1993_production}.
This symbolic capital might provide opportunities to acquire material advantages such as collaborations with other institutions and access to broader audiences, ultimately bolstering the museum's longevity.

While limited research exists that investigates reputation effects for (art) museums, its effects have been investigated for non-profits more widely.
Here reputation is often referred to as an organization's resource or asset \parencite{Kong_Farrell_2010_image,Goldfarb_Zavyalova_Pillai_2018_Rao,Walker_McCarthy_2010_survival,Schloderer_Sarstedt_Ringle_2014_reputation}. 
Studies primarily focus on the effect of reputation on some tangible benefits such as donations and contributions \parencite{Sarstedt_2010_reputation,Mews_Boenigk_2012_blood,Heller_2008_alliances,Meijer_2009_reputation}, or its direct effect on survial \parencite{Lu_Shon_Zhang_2019_dissolution,Singh_1991_change,Hager_2001_vulnerability,Bielefeld_1994_survival,HernandezOrtiz_2022_discontinuity}\footnote{Furthermore, the term "survival" is sometimes used more loosely to refer to a non-profit's effective operation and financial condition rather than the more narrowly defined dissolution (e.g. \cite[p.58]{Santos_Laureano_Moro_2019_reputation}, \cite[p.15]{Heller_2008_alliances}, \cite[p.111]{Schloderer_Sarstedt_Ringle_2014_reputation})}, with most studies finding support for a positive effect of reputation.
Next to the reputation of the general public, studies have also found reputation effects among more narrowly defined groups, in particular peers \parencite{Padanyi_Gainer_2003_peerreputation} and networks of similar organizations on regional and national scale \parencite{Walker_McCarthy_2010_survival}.
In particular this reputation among more narrowly defined groups makes it likely that the reputation of the  \emph{founder} matters (beyond the reputation of the \emph{organization}, the focus on most prior research) given the highly personal character of the private museum initiative and the central role of the founder in negotiations with decision makers from other organizations or the public sector.

In the case of the founder losing his reputation, the consequences then might extend beyond personal demotivation.
Such a decline could also curtail potential opportunities for cooperation with other institutions and limit the museum's reach to larger audiences, thus posing a risk to the museum's continued operation and survival in the long term.
\bigbreak
\noindent
\textbf{Hypothesis 1b}: Private museums are more likely to close after a decline of their founder's status.

\paragraph*{Identity}

The perception of the museum itself, beyond the reputation of the founder, might also affect survival prospects.
An identity that is legitimate (understandable) and favorable (positive) has been argued to have a vast number of positive outcomes for organizations such as allowing higher profits, satisfaction, as well as weaker impact of shocks, and providing the opportunity to deviate from established patterns  \parencite{Lange_Lee_Dai_2010_reputation}, which in turn can facilitate survival prospects \parencite{Rao_1994_reputation}.
In the context of private museums, a more positive identity could lead to higher visitor numbers, discounts from art dealers, government grants or collaborations with other museums or corporations, resulting in higher revenues and lower acquisition costs.

@@

It is however necessary to consider that in particular the name of a museum might be associated with a museum's survival via identity, but also reflect founder intentions regarding the institutions longevity; this possibility is discussed in more detail in the results section.
\paragraph*{Competition}

Just as the environment can be supportive of museums recognized perceived as renowned, it can also be a source of risk if it such support is absent.
In particular this might be the case under conditions of harsh competition, which private museums have been characterized to engage in.
While some self-descriptions by founders focus on collaborative aspects \parencite{Duron_2020_rebaudengo,BMW_2016_salsali}, previous studies have highlighted how private museums compete with public museums over art works, audiences, sponsoring and donors \parencite[p.4]{Kolbe_etal_2022_privatemuseum}. 
So far the primary interest of investigating competition has been to evaluate whether private museums pose a threat to art provision by public institutions.
However, less focus has been given to the fact that operating in a competitive environment may also have consequences for the private museums themselves, in particular their survival aspects.

While private museums presumably depend on their founder to a large extent, income from tickets and facilities, as well sponsorships and public support possibly also constitute other non-negligible revenue streams.
Competition directly has not yet been investigated as a cause for private museum closing, but a possible mechanism of competition, "insufficient interest from the public", has been identified as the second most reported closure reason by \textcite[p.6]{Velthuis_Gera_2024_fragility}. 
Lack of interest by the public can result from multiple processes.
While competition is one such processes, and applies if nearby competing institutions capture large portions of the audience, lack of interest can also result from a lack of audience (in the absence of competing institutions), which may be the case for private museums located in sparsely populated areas.
In both these cases, the resources that a private museum is able to receive from actors in its environment are limited.

\textbf{Hypothesis 3}: A private museum is more likely to close in a more competitive environment.
\subsection*{Methods and Data}

A newly constructed database, assembled from a wide range of sources, is used to measure the survival of private museums (cf. \textcite{Velthuis_etal_2023_boom} for details).
Currently, the database includes 550 museums, of which 453 are currently open, 72 have closed and 25 have transformed into other organizational forms.\footnote{The 25 institutions which were private museums at some point and have since been transformed into other organizational forms are not included in this analysis and discussed in the limitations.}
Using this database, in particular year of opening and, if available, closing, it is possible to reconstruct the life course of each institution with museum-year as the unit of analysis.
While the database is not limited to private museums of a particular time period, due to the recency of the private museum boom in the last two decades 80\% of the museums included in it have opened after 2000.
As the database was constructed retrospectively from 2020 to 2022, I focus here on closing events from 2000 to 2021 to limit the possibility that some institutions were missed, as such risks are likely higher for earlier time periods.
Furthermore, a small number of museums are excluded, in particular 2 museums for which the closing date could not be found, 17 museums which opened in 2021 and later (as survival analyses requires positive survival times), 1 museum the opening year of which could not be determined, as well as 2 museums which closed before 2000 and thereby fall outside the period under investigation.
This leaves 503 private museums at risk which have accumulated 6096 museum-years and 68 closures.
This data allows to conduct survival analysis (also known as event history analysis), the most common statistical technique to investigate hazard rates and identify mortality-assocated covariates \parencite{Moore_2015_survival,Allison_2014_event}.
